\myTitle

### 2024-05-13 body-text-of-paper.tex
Disable endfloat for one-column version
To send to Tony

%%%%%%%%%%%
\begin{figure}
\newcommand{\gspcapone}{\footnotesize
A genomic simulation pedigree (GSP) for simulating $F_2$ individuals.
Squares denote founder and non-founder nodes in the pedigree. There are four founders in the pedigree,
individuals 1--4.  The  triangles above each of those founders denotes, in color
and text, the population of origin of each of the two genomes within each founder.
Thus, individuals~1 and~2 are founders purely from population~$A$, and~3 and~4
are purely from population~$B$.  The pink hexagon labeled s7 is the sample node.
It represents samples that are taken from non-founder node~7, which, from the structure
of the pedigree, clearly represents an $F_2$ hybrid between populations $A$ and $B$.  The red numbers
on edges above and leading into each non-founder node indicate the number of
gametes that get segregated from the parent node (at the top of the edge) to the
daughter node (at the bottom of the edge).
This emphasizes that,
in dropping genes through a GSP, in contrast
to a traditional pedigree, each parent
node may segregate multiple gametes to a daughter node.  See the text for
further explanation.
Each non-founder node $x$ will have exactly two edges coming into it from above
(from its two parents), and we denote the numbers associated with those edges as $G^+_{x,1}$ and $G^+_{x,2}$.
For example, node 6 has $G^+_{6,1}=G^+_{6,2} = 2$.
A founder individual may have, at most, 2 edges coming downward out of it, while a non-founder individual may have more or less than two edges coming downward
out of it.  Letting $e_x$ be the number of downward edges out of node $x$ to its $e_x$
non-founder daughter nodes, we denote the number of gametes segregated
through each edge as $G^{-}_{x, 1},\ldots,G^{-}_{x, e_x}$.  For example, at
non-founder node~5, $e_5 = 1$ and $G^{-}_{5,1}=4$.
Finally, the purple number adjacent to the edge leading into the
sample node indicates how many sample individuals are drawn from individual
node 7. Any non-founder individual node can have a maximum of one edge leading
to a sample node.  We denote the number of samples emanating from node $x$ as $S_x$. In our example $S_7=4$ indicates that four $F_2$ individuals are
created from a single simulation on this GSP.
}
\begin{center}
\includegraphics[width=\columnwidth]{figures/F2.pdf}
\end{center}
\caption[\gspcapone]{\gspcapone}
\label{fig:gsp1}
\end{figure}
%%%%%%%%%%%%
%%%%%
\begin{figure}
\newcommand{\gspcaptwo}{\footnotesize
An example GSP generated by {\footnotesize\tt create\_GSP()} that produces four different
admixture categories: one~$F_1$ in sample node~s7; two~$F_2$'s in sample node~s11; one $\BC_1$ in sample node~s9; and two $\BC_2$'s in sample node~s10.
}
\begin{center}
\includegraphics[width=\columnwidth]{figures/AllThem.pdf}
\end{center}
\caption[\gspcaptwo]{\gspcaptwo}
\label{fig:gsp2}
\end{figure}
%%%%

@@

We performed 500 replicate simulations with \gscramble{} using either 30 or 65 markers.  Each
simulated data set was then analyzed using {\sc NewHybrids} \citep{anderson2002model}. Individuals not used as founders were given the ``s'' and ``z'' flags, indicating that they were considered training samples.  The remaining 46 individuals were treated as possible hybrids and their genetic data used to calculate the posterior probability that they belonged to one of 10 hybrid categories as shown in Table~\ref{tab:newhybcats}, using 25,000 MCMC sweeps and discarding the first 5,000 as burn-in.
%%%%%%%%%%%%
\begin{table}
\caption{The 10 hybrid categories considered for inference with {\sc NewHybrids} and the
expected frequency of their genotypes having 2 ($f(C,C)$), 1 ($f(C,S)$) or 0 ($f(S,S)$) copies of ancestry from cutthroat trout.}
\label{tab:newhybcats}
{\small
\begin{tabular*}{\columnwidth}{@{\extracolsep{\fill}} lrrr}
\hline\hline
Category	& $f(C,C)$ &	$f(C,S)$ & 	$f(S,S)$ \\ \hline
\input{inputs/newhybs-cats.tsv}
\end{tabular*}
}
\vspace*{-2.3ex}\hrule\vspace*{0.3ex}\hrule
\end{table}
%%%%%%%%%%%%

@@

When hybrids were simulated in a way that accounts for the physical linkage of
markers, the variance in admixture fractions was considerably higher than when
markers were simulated under the assumption that none of them are physically
linked (Fig.~\ref{fig:ad-fract-sims}).
%%%%%%%%%%
\begin{figure}
\newcommand{\adfcap}{\footnotesize Admixture fractions of individuals simulated
assuming no physical linkage (top four panels)
and when physical linkage is simulated using \gscramble{} assuming a genetic map with 18 chromosomes spanning 2.35 gigabases with recombination fractions like those found in pigs
\citep{tortereau2012high}.  In the top 4 panels the number
of simulated loci appears in the panel header.  Failing to simulate markers with physical linkage
produces distributions of admixture fractions that are far more narrow than reality, giving the
mistaken impression that admixture fractions of different hybrid categories do not overlap.  }
\includegraphics[width=\columnwidth]{figures/link-and-nolink-sims-crop.pdf}
\caption[\adfcap]{\adfcap}
\label{fig:ad-fract-sims}
\end{figure}
%%%%%%%%%%
If markers are assumed to be unlinked, then the variance around the $Q$ estimates of different 
hybrid categories continually declines as more markers are added, from 100, to 100,000, which 
makes it appear that different hybrid categories become increasingly well-resolved.  For example, 
looking at the distributions for 100,000 unlinked markers one would think that it would be easy to 
distinguish between the $BC_1$ through $BC_4$ categories on the basis of the admixture fractions.
This is, however, not the case.  The bottom panel of Fig.~\ref{fig:ad-fract-sims} shows the distribution
of the true admixture fraction (calculated as the total length of genome from each of the two
different species) of each individual when simulated with physical linkage using \gscramble{}.
When physical linkage is not ignored, it is quite clear that there is considerable overlap in the
admixture fractions between the $BC_1$ and $BC_2$ categories, and that $BC_2$, $BC_3$, and 
$BC_4$ all have admixture-fraction distributions that overlap greatly.  This shows that,
in real life, when physical linkage cannot be ignored, it may be quite difficult to distinguish
between different backcross categories on the basis of their admixture fractions.

@@

As expected, there was no observed RISPI in the ADMIXTURE-estimated
ancestry proportions  when simulating individuals
using \gscramble{} (Fig.~\ref{fig:norispi}).  On average, the estimated $Q$ values
were around 0.5, which is expected when the samples are drawn from the same
population.
%%%
\begin{figure}
\newcommand{\gssimcap}{\footnotesize ADMIXTURE estimates of ancestry
fractions, $Q$, for individuals simulated from samples from the same population,
simulated using \gscramble{}.  There is considerable variation in the estimates,
but the central tendency of them all is around 0.5, as would be expected in this case
within which the different
samples are drawn from the same population.  Contrast this to the corresponding panel ($L=1,000, N=50$) in Fig.~\ref{fig:bias-sims}(b),
in which the RISPI is readily apparent.}
\includegraphics[width=\columnwidth]{figures/gscram_supervised_Qs.pdf}
\caption[\gssimcap]{\gssimcap}
\label{fig:norispi}
\end{figure}
%%%

@@

The barplots in Fig.~\ref{fig:newhyb-sims-barplot} show the distribution of posterior
probabilities of the different hybrid categories. 
%%%%%%%%%%
\begin{figure}
\newcommand{\nhbpcap}{\footnotesize Posterior probabilities for simulated
individuals of different trout hybrid categories.  Each pair of plots shows the results
for individuals of the true hybrid category listed atop the pair  (see
 Table~\protect\ref{tab:newhybcats} for the meanings of the names).  The top panel of each pair
shows results for fish simulated with the 30 markers used by \protect\citet{rizza2023limited}, and the bottom panel shows results for fish simulated with the original 30 markers,
plus an additional  35 (65 total), some of which are linked
on the same chromosome. Each individual appears as a thin vertical line, posterior probabilities of
each inferred hybrid category are indicated by extent of different colors on those lines.  Individuals
are sorted according to maximum {\em a posteriori} hybrid category, and, within that by the
maximum posterior probability. }
{\centering
\includegraphics[width=0.98\columnwidth]{figures/newhybs-texed-crop.pdf}
}
\caption[\nhbpcap]{\nhbpcap}
\label{fig:newhyb-sims-barplot}
\end{figure}
%%%%%%%%%%
These are similar to STRUCTURE or ADMIXTURE barplots;
however, rather than plotting the ancestry fraction, the different colors are showing the
posterior probabilities of different hybrid categories.  Each individual is represented by a thin
vertical line, and the individuals have been arranged from left to right and the colors have been
arranged from top to bottom throughout different sections of the plot to make it easy to see
the patterns.  The {\em maximum-a-posteriori} (MAP) hybrid category for any individual is represented
by the color that is on the bottom of the bar, so it is easy to see the shifts where the MAP hybrid
category changes from one to another.  It is apparent over all individuals simulated to be
either CCT (cutthroat trout), $F_1$, $F_2$, $BC_1$--CCT, or $BC_2$--CCT, that the MAP estimate of their
hybrid categories were correct for largely the same number of individuals whether 30 or 65 markers
were used. By contrast, for the individuals simulated to be of the complex hybrid category
$BC_1$--CCT~$\times$~$BC_2$--CCT, the MAP estimate from 65 markers was correct more often 
than from 
30 markers.  However, it is also apparent that using 65 markers, some of which are physically
linked, in the software {\em NewHybrids}---which assumes that all markers are unlinked---yields 
estimates of uncertainty that are inaccurate.  Tellingly, when the estimates using 65 markers are 
incorrect (for example, when $BC_1$--CCT individuals have a maximum posterior of being in the 
$BC_2$--CCT category,
or when individuals that are truly of the $BC_1$--CCT~$\times$~$BC_2$--CCT category have a
maximum posterior of being from the $BC_2$--CCT category), many of those incorrect posteriors
are quite high ($>0.90$).  In fact, using 65 markers yields more incorrect {\em and high} posteriors 
than does using the 30 unlinked markers. So, as expected, ignoring physical linkage (e.g., as 
implemented in {\sc NewHybrids}) 
can yield a poor estimate of the uncertainty around different inferences.

@@

%%%
\begin{figure}
\newcommand{\pigsimsqscap}{\footnotesize (A) The density of the estimated Q values from ADMIXTURE for
the simulated runs are
grouped across all population-combinations for each hybrid class and colored by the $K$ genetic cluster
(e.g., the Pop1 density plot for the $\BC_1$ plots the Pop1 genetic cluster for both
the Pop1-Pop2 and Pop1-Pop3 combinations). The density plots across the $K$
genetic clusters are mostly similar with the most variation among the genetic clusters evident in
the $F_1$ plot. (B) Plotting the density of the estimated Q values from ADMIXTURE for the $F_1$ hybrids,
separated by the population combinations, reveals the least variance in estimated Q values for
Pop1-Pop2 compared to Pop1-Pop3 and Pop2-Pop3. Note that these population combinations group the
reciprocal population combination in each plot (e.g., "Pop1\_Pop2" has simulations
from both Pop1-Pop2 and Pop2-Pop1).}
\includegraphics[width=\columnwidth]{figures/simulation-density-multipanel.pdf}
\caption[\pigsimsqscap]{\pigsimsqscap}
\label{fig:pig-sim-qs}
\end{figure}
%%%

>>>>

%%%%%%%%%%%
\begin{figure}
\newcommand{\gspcapone}{\footnotesize
A genomic simulation pedigree (GSP) for simulating $F_2$ individuals.
Squares denote founder and non-founder nodes in the pedigree. There are four founders in the pedigree,
individuals 1--4.  The  triangles above each of those founders denotes, in color
and text, the population of origin of each of the two genomes within each founder.
Thus, individuals~1 and~2 are founders purely from population~$A$, and~3 and~4
are purely from population~$B$.  The pink hexagon labeled s7 is the sample node.
It represents samples that are taken from non-founder node~7, which, from the structure
of the pedigree, clearly represents an $F_2$ hybrid between populations $A$ and $B$.  The red numbers
on edges above and leading into each non-founder node indicate the number of
gametes that get segregated from the parent node (at the top of the edge) to the
daughter node (at the bottom of the edge).
This emphasizes that,
in dropping genes through a GSP, in contrast
to a traditional pedigree, each parent
node may segregate multiple gametes to a daughter node.  See the text for
further explanation.
Each non-founder node $x$ will have exactly two edges coming into it from above
(from its two parents), and we denote the numbers associated with those edges as $G^+_{x,1}$ and $G^+_{x,2}$.
For example, node 6 has $G^+_{6,1}=G^+_{6,2} = 2$.
A founder individual may have, at most, 2 edges coming downward out of it, while a non-founder individual may have more or less than two edges coming downward
out of it.  Letting $e_x$ be the number of downward edges out of node $x$ to its $e_x$
non-founder daughter nodes, we denote the number of gametes segregated
through each edge as $G^{-}_{x, 1},\ldots,G^{-}_{x, e_x}$.  For example, at
non-founder node~5, $e_5 = 1$ and $G^{-}_{5,1}=4$.
Finally, the purple number adjacent to the edge leading into the
sample node indicates how many sample individuals are drawn from individual
node 7. Any non-founder individual node can have a maximum of one edge leading
to a sample node.  We denote the number of samples emanating from node $x$ as $S_x$. In our example $S_7=4$ indicates that four $F_2$ individuals are
created from a single simulation on this GSP.
}
\begin{center}
\includegraphics[width=0.48\textwidth]{figures/F2.pdf}
\end{center}
\caption[\gspcapone]{\gspcapone}
\label{fig:gsp1}
\end{figure}
%%%%%%%%%%%%
%%%%%
\begin{figure}
\newcommand{\gspcaptwo}{\footnotesize
An example GSP generated by {\footnotesize\tt create\_GSP()} that produces four different
admixture categories: one~$F_1$ in sample node~s7; two~$F_2$'s in sample node~s11; one $\BC_1$ in sample node~s9; and two $\BC_2$'s in sample node~s10.
}
\begin{center}
\includegraphics[width=0.48\textwidth]{figures/AllThem.pdf}
\end{center}
\caption[\gspcaptwo]{\gspcaptwo}
\label{fig:gsp2}
\end{figure}
%%%%

@@

We performed 500 replicate simulations with \gscramble{} using either 30 or 65 markers.  Each
simulated data set was then analyzed using {\sc NewHybrids} \citep{anderson2002model}. Individuals not used as founders were given the ``s'' and ``z'' flags, indicating that they were considered training samples.  The remaining 46 individuals were treated as possible hybrids and their genetic data used to calculate the posterior probability that they belonged to one of 10 hybrid categories as shown in Table~\ref{tab:newhybcats}, using 25,000 MCMC sweeps and discarding the first 5,000 as burn-in.
%%%%%%%%%%%%
\begin{table}
\caption{The 10 hybrid categories considered for inference with {\sc NewHybrids} and the
expected frequency of their genotypes having 2 ($f(C,C)$), 1 ($f(C,S)$) or 0 ($f(S,S)$) copies of ancestry from cutthroat trout.}
\label{tab:newhybcats}
{\small
\begin{tabular*}{0.48\textwidth}{@{\extracolsep{\fill}} lrrr}
\hline\hline
Category	& $f(C,C)$ &	$f(C,S)$ & 	$f(S,S)$ \\ \hline
\input{inputs/newhybs-cats.tsv}
\end{tabular*}
}
\vspace*{-2.3ex}\hrule\vspace*{0.3ex}\hrule
\end{table}
%%%%%%%%%%%%

@@

When hybrids were simulated in a way that accounts for the physical linkage of
markers, the variance in admixture fractions was considerably higher than when
markers were simulated under the assumption that none of them are physically
linked (Fig.~\ref{fig:ad-fract-sims}).
%%%%%%%%%%
\begin{figure}
\newcommand{\adfcap}{\footnotesize Admixture fractions of individuals simulated
assuming no physical linkage (top four panels)
and when physical linkage is simulated using \gscramble{} assuming a genetic map with 18 chromosomes spanning 2.35 gigabases with recombination fractions like those found in pigs
\citep{tortereau2012high}.  In the top 4 panels the number
of simulated loci appears in the panel header.  Failing to simulate markers with physical linkage
produces distributions of admixture fractions that are far more narrow than reality, giving the
mistaken impression that admixture fractions of different hybrid categories do not overlap.  }
\includegraphics[width=0.48\textwidth]{figures/link-and-nolink-sims-crop.pdf}
\caption[\adfcap]{\adfcap}
\label{fig:ad-fract-sims}
\end{figure}
%%%%%%%%%%
If markers are assumed to be unlinked, then the variance around the $Q$ estimates of different 
hybrid categories continually declines as more markers are added, from 100, to 100,000, which 
makes it appear that different hybrid categories become increasingly well-resolved.  For example, 
looking at the distributions for 100,000 unlinked markers one would think that it would be easy to 
distinguish between the $BC_1$ through $BC_4$ categories on the basis of the admixture fractions.
This is, however, not the case.  The bottom panel of Fig.~\ref{fig:ad-fract-sims} shows the distribution
of the true admixture fraction (calculated as the total length of genome from each of the two
different species) of each individual when simulated with physical linkage using \gscramble{}.
When physical linkage is not ignored, it is quite clear that there is considerable overlap in the
admixture fractions between the $BC_1$ and $BC_2$ categories, and that $BC_2$, $BC_3$, and 
$BC_4$ all have admixture-fraction distributions that overlap greatly.  This shows that,
in real life, when physical linkage cannot be ignored, it may be quite difficult to distinguish
between different backcross categories on the basis of their admixture fractions.

@@

As expected, there was no observed RISPI in the ADMIXTURE-estimated
ancestry proportions  when simulating individuals
using \gscramble{} (Fig.~\ref{fig:norispi}).  On average, the estimated $Q$ values
were around 0.5, which is expected when the samples are drawn from the same
population.
%%%
\begin{figure}
\newcommand{\gssimcap}{\footnotesize ADMIXTURE estimates of ancestry
fractions, $Q$, for individuals simulated from samples from the same population,
simulated using \gscramble{}.  There is considerable variation in the estimates,
but the central tendency of them all is around 0.5, as would be expected in this case
within which the different
samples are drawn from the same population.  Contrast this to the corresponding panel ($L=1,000, N=50$) in Fig.~\ref{fig:bias-sims}(b),
in which the RISPI is readily apparent.}
\includegraphics[width=0.48\textwidth]{figures/gscram_supervised_Qs.pdf}
\caption[\gssimcap]{\gssimcap}
\label{fig:norispi}
\end{figure}
%%%

@@

The barplots in Fig.~\ref{fig:newhyb-sims-barplot} show the distribution of posterior
probabilities of the different hybrid categories. 
%%%%%%%%%%
\begin{figure}
\newcommand{\nhbpcap}{\footnotesize Posterior probabilities for simulated
individuals of different trout hybrid categories.  Each pair of plots shows the results
for individuals of the true hybrid category listed atop the pair  (see
 Table~\protect\ref{tab:newhybcats} for the meanings of the names).  The top panel of each pair
shows results for fish simulated with the 30 markers used by \protect\citet{rizza2023limited}, and the bottom panel shows results for fish simulated with the original 30 markers,
plus an additional  35 (65 total), some of which are linked
on the same chromosome. Each individual appears as a thin vertical line, posterior probabilities of
each inferred hybrid category are indicated by extent of different colors on those lines.  Individuals
are sorted according to maximum {\em a posteriori} hybrid category, and, within that by the
maximum posterior probability. }
{\centering
\includegraphics[width=0.48\textwidth]{figures/newhybs-texed-crop.pdf}
}
\caption[\nhbpcap]{\nhbpcap}
\label{fig:newhyb-sims-barplot}
\end{figure}
%%%%%%%%%%
These are similar to STRUCTURE or ADMIXTURE barplots;
however, rather than plotting the ancestry fraction, the different colors are showing the
posterior probabilities of different hybrid categories.  Each individual is represented by a thin
vertical line, and the individuals have been arranged from left to right and the colors have been
arranged from top to bottom throughout different sections of the plot to make it easy to see
the patterns.  The {\em maximum-a-posteriori} (MAP) hybrid category for any individual is represented
by the color that is on the bottom of the bar, so it is easy to see the shifts where the MAP hybrid
category changes from one to another.  It is apparent over all individuals simulated to be
either CCT (cutthroat trout), $F_1$, $F_2$, $BC_1$--CCT, or $BC_2$--CCT, that the MAP estimate of their
hybrid categories were correct for largely the same number of individuals whether 30 or 65 markers
were used. By contrast, for the individuals simulated to be of the complex hybrid category
$BC_1$--CCT~$\times$~$BC_2$--CCT, the MAP estimate from 65 markers was correct more often 
than from 
30 markers.  However, it is also apparent that using 65 markers, some of which are physically
linked, in the software {\em NewHybrids}---which assumes that all markers are unlinked---yields 
estimates of uncertainty that are inaccurate.  Tellingly, when the estimates using 65 markers are 
incorrect (for example, when $BC_1$--CCT individuals have a maximum posterior of being in the 
$BC_2$--CCT category,
or when individuals that are truly of the $BC_1$--CCT~$\times$~$BC_2$--CCT category have a
maximum posterior of being from the $BC_2$--CCT category), many of those incorrect posteriors
are quite high ($>0.90$).  In fact, using 65 markers yields more incorrect {\em and high} posteriors 
than does using the 30 unlinked markers. So, as expected, ignoring physical linkage (e.g., as 
implemented in {\sc NewHybrids}) 
can yield a poor estimate of the uncertainty around different inferences.

@@

%%%
\begin{figure}
\newcommand{\pigsimsqscap}{\footnotesize (A) The density of the estimated Q values from ADMIXTURE for
the simulated runs are
grouped across all population-combinations for each hybrid class and colored by the $K$ genetic cluster
(e.g., the Pop1 density plot for the $\BC_1$ plots the Pop1 genetic cluster for both
the Pop1-Pop2 and Pop1-Pop3 combinations). The density plots across the $K$
genetic clusters are mostly similar with the most variation among the genetic clusters evident in
the $F_1$ plot. (B) Plotting the density of the estimated Q values from ADMIXTURE for the $F_1$ hybrids,
separated by the population combinations, reveals the least variance in estimated Q values for
Pop1-Pop2 compared to Pop1-Pop3 and Pop2-Pop3. Note that these population combinations group the
reciprocal population combination in each plot (e.g., "Pop1\_Pop2" has simulations
from both Pop1-Pop2 and Pop2-Pop1).}
\includegraphics[width=0.48\textwidth]{figures/simulation-density-multipanel.pdf}
\caption[\pigsimsqscap]{\pigsimsqscap}
\label{fig:pig-sim-qs}
\end{figure}
%%%
